Copyright

### 2025-01-26 choose_license.qmd
Various minor improvements

So far, you took care to legally include works by others in your project folder. Up next, you will free up your project for reuse. This means ensuring that every bit (co-)authored by you in the project is put under a free/open license. It may be that your project (i.e., its files and folders) is not one work, but rather consists of multiple works available under different licenses. In that case, you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project). Sometimes, even different parts of a file might be subject to different licenses. For the purpose of this tutorial, we will consider the manuscript which you edited and the data dictionary you created.

@@

We recommend to choose free/open licenses suitable to the type of work (i.e., code, data, and everything else). [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) are good choices to apply simultaneously to all own works. Take your time to read the summaries behind the links and/or the actual license text to understand the legal effect. You can adapt the following wording to your use case:

@@

Finally, there may be some minor files which are build artifacts. You can either add them to your `.gitignore` file or use the CC0\ 1.0 license/waiver with a copyright tag such as `SPDX-FileCopyrightText: NONE` to assert that there is no copyright holder. For more information, also discussing other corner cases, you can read their [Frequently Asked Questions](https://reuse.software/faq/).

>>>>

So far, you took care to legally include works by others in your project folder. Up next, you will free up your project for reuse. This means ensuring that every bit (co-)authored by you in the project is put under a free/open license. It may be that your project (i.e., its files and folders) is not one work, but rather consists of multiple works available under different licenses. In that case, you need to indicate the license on a per-file or per-folder basis (rather than choosing one for the whole project). Sometimes, even different parts of a file might be subject to different licenses. For the purpose of this tutorial, we will consider the manuscript and the bibliography which you edited as well as the data dictionary you created.

@@

We recommend to choose free/open licenses suitable to the type of work (e.g., code, data, image etc.). [CC0\ 1.0](https://creativecommons.org/publicdomain/zero/1.0/) and [Apache\ 2.0](https://choosealicense.com/licenses/apache-2.0/) are good choices to apply simultaneously to all own works. Take your time to read the summaries behind the links and/or the actual license text to understand the legal effect. You can adapt the following wording to your use case:

@@

Finally, there may be some minor files which are build artifacts. You can either add them to your `.gitignore` file or use CC0\ 1.0 with a copyright tag such as `SPDX-FileCopyrightText: NONE` to assert that there is no copyright holder. For more information, also discussing other corner cases, you can read their [Frequently Asked Questions](https://reuse.software/faq/).

### 2025-01-26 copyright.qmd
Add full license texts
Various minor improvements

Licenses may either require that one mentions the applicable license by name and/or link ("Indicate license") or to even copy the full license text along with the used work ("Add license text"). Of course, the latter may not be feasible if the work is printed and smaller than the license text. For example, if the license of an image requires to print the full license text, the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].

@@

If you modified the work and put a sufficient amount of creativity into it, the result is called a _derivative work_. It is (again) protected by copyright and even if you share it, nobody except you is allowed to copy, modify, or distribute it further. You can put it under a free/open license but don't have to. The only exception is if the original work has been made available under a "copyleft" license. In that case, if you share the work with others you must put it under the same license as the original work^[or sometimes also under a later version or under a compatible license] -- or you commit a copyright violation.^[Note, however, that [copyleft licenses are not automatic](https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic): A derivative work of a copylefted work is not "implicitly" under the same license -- others would commit a copyright violation if they just took it. Only if the derivative author explicitly makes it available under the same license, others may use the derivative work under the same terms.] This explains the origin of the term _copyleft_, as it "turns copyright around" to make works permanently free: ©\ →\ 🄯. In the realm of Creative Commons licenses, copyleft is also called _share alike_. The opposite of a copyleft license is sometimes also called a _permissive_ license.

@@

If the license required adding the full license text, you would also need to copy it to `LICENSE.txt`.

>>>>

Licenses may either require that one mentions the applicable license by name and possibly link ("Indicate license") or they even require to copy the full license text along with the used work ("Add license text"). Of course, the latter may not be feasible if the work is printed and smaller than the license text. For example, if the license of an image requires to print the full license text, the image mostly cannot be used in newspapers [see @Wikipedia2024FDL].

@@

If you modified the work and put a sufficient amount of creativity into it, the result is called a _derivative work_. It is (again) protected by copyright and even if you share it, nobody except you is allowed to copy, modify, or distribute it further. You can put it under a free/open license but don't have to.^[Of course, you still need to abide by the license conditions of the original work, such as attribution.] The only exception is if the original work has been made available under a "copyleft" license. In that case, if you share the work with others you must put it under the same license as the original work^[or sometimes also under a later version or under a compatible license] -- or you commit a copyright violation.^[Note, however, that [copyleft licenses are not automatic](https://commons.wikimedia.org/wiki/Commons:Viral_licenses_are_not_automatic): A derivative work of a copylefted work is not "implicitly" under the same license -- others would commit a copyright violation if they just took it. Only if the derivative author explicitly makes it available under the same license, others may use the derivative work under the same terms.] This explains the origin of the term _copyleft_, as it "turns copyright around" to make works permanently free: ©\ →\ 🄯. In the realm of Creative Commons licenses, copyleft is also called _share alike_. The opposite of a copyleft license is sometimes also called a _permissive_ license.

@@

If the license required adding the full license text, you would also need to copy it to the project folder.
